Type Error Telemetry at Scale

### 2023-09-29 programming23/paper.tex
sec 3 edits

Scripts come in two flavors: \emph{module scripts} and \emph{non-module
scripts}.
Module scripts provide reusable libraries, which may be
{imported} by other scripts.
Module scripts therefore form a graph in general,
though it is an error in \mstrict{} mode to create an import cycle
and edges are removed to make the graph acyclic.
%% \QALAN{}: do all modes remove the edges?

@@

\subsection{Limits of Telemetry}

Telemetry allows an application to ``phone home'' with data
summarizing usage patterns, such performance data, crash reporting,
and feature uptake. A typical usage is in deciding whether an API can
be deprecated: without telemetry it may be difficult to know how
popular it is, but with telemetry it is
straightforward.

IDEs use telemetry in a similar fashion\todo{I really don't know what
  this means. Are you saying VSCode is constantly calling home? Is it
  telling Racket about my API usage? Do we need anything in this para?} to most user-facing
applications, for example VSCode~\cite{vsc-telemetry} and
IntelliJ~\cite{intellij-telemetry} report telemetry.
See~\cref{s:related} for futher comparisons among telemetry designs.

Telemetry for programming languages can be controversial:e.g., see
the lively discussion around telemetry in the Go
toolchain~\cite{golang-telemetry}. 
The Luau open source toolchain does \emph{not} report telemetry,
as it is designed to be used in build environments such as Continuous Integration
servers, where hermetic deterministic builds are expected.

Telemetry therefore faces at least two major constraints:
\begin{itemize}

\item It should not reveal private information. This can include
  Personally Identifying
  Information (PII) about creators: their identity, location, etc. It also
  includes trade secrets. Even an error message that contains the name
  of a function could reveal something the creator would not want to
  share.

\item For performance reasons, it should have minimal run-time
  overhead and also transmit small and bounded amounts of data.
\end{itemize}
Cox discusses in detail the privacy implications and tradeoffs
in telemetry~\cite{transparent-telemetry}.

% BEN: I believe the above subsumes all of the below:
% Much of the concern about telemetry is because of its potential impact on
% privacy. It can not only reveal Personally Identifiable Information
% (PII) but also trade secrets, etc. Thus, telemetry should not reveal
% PII, source code, error messages (which can refer to source), the
% creator's identity, and so on.
% % and no information about what creation the data came from.

% For performance reasons, there are limits to telemetry data:
% telemetry records have size limits, and the performance impact of
% recording telemetry should be minimal.

% In summary, the limitations of telemetry are:
% \begin{itemize}
%   \item Data that don't reveal private information.
%   \item Fixed-size telemetry records.
%   \item Minimal overhead to record telemtry.
% \end{itemize}

In addition, due to the architecture of Roblox Studio
there is some information that is not available to the
typechecker:
\begin{itemize}
  \item The typechecker does not have access to some lifecycle events,
    such as save, quit and publish.
  \item The typechecker does not have access to GUI state, for example
    whether the Script Analysis widget is visible.
\end{itemize}

@@

\begin{figure}[t]\centering
  % BEN: The left brace is confusing. Is ID not fixed?
  \begin{tabular}{|l|l|l|l|l|l|l|}\hline
    ID & Time & Mode & Reason & Size Info & Overall Counts & Edit Range Counts
    \\\toprule
    \multicolumn{6}{c}{\upbracefill} & \multicolumn{1}{c}{\upbracefill} \\[-0.5ex]
    \multicolumn{6}{c}{fixed-size} & \multicolumn{1}{c}{variable-size}
  \end{tabular}
  \caption{Structure of a telemetry record. There are 13 overall counts and up to 70 edit range counts in each record.}
  \label{f:tdata}
\end{figure}

The telemetry for a Roblox Studio session is gathered as a series of
records, all with the same \panon{} session
identifier. This allows us to correlate telemetry across a single
session, but not between sessions.

In order to avoid swamping our telemetry servers, we randomly throttled
the generation of telemetry records:
\begin{itemize}
  \item
    1\% of Roblox Studio sessions were enrolled in generating telemetry, and
  \item
    0.5\% of runs of the typechecker (approximately 1 in 200 keystrokes)
      in an enrolled session generated a telemetry record.
\end{itemize}
In order to find how file-switching correlated with type errors, we
additionally recorded
every run of the typechecker in an enrolled session which switched
file generated a telemetry record.\todo{I cannot parse this!}

@@

To track the edit range, we record a start and end position, which we
update appropriately on every edit. This can result in very large edit
ranges, for example, if the user edits at the beginning and end of the
file.
The main benefit of this strategy is that it ensures constant-size telemetry records.

\paragraph{NOTES} %% FILL
This kind of telemetry data is coarse-grained.
For instance, it does not let us distinguish between edits that ignore a type
error from edits that remove one error while introducing another.
We acknowledge this and other threats in~\cref{s:threats}.
The main advantages of the telemetry design are the small size of each record
and complete the lack of PII.
Furthermore, despite the limitations, Studio's telemetry supports a variety of
interesting conclusions that we showcase in the next section.

@@

@@

@@

In other professional-grade IDEs, such as VSCode~\cite{vscode-telemetry}
and \code{.NET}~\cite{dotnet-telemetry}, telemetry may include all sorts of
data: error reports, source code fragments, timestamps, and filepaths.
Users may be able to opt out of some telemetry, but the details depend on the
license agreement.
Furthermore, at least in VSCode, IDE extensions can report their own telemetry.
While this data can be invaluable for discovering bugs in production, it
must be handled with extreme care.

>>>>

Scripts come in two flavors: \emph{module scripts} and \emph{non-module
scripts}.
(Though, throughout this paper we use the word ``module''
to refer to arbitrary scripts.)
Module scripts provide reusable libraries, which may be
{imported} by other scripts.
Module scripts therefore form a graph in general,
though it is an error in \mstrict{} mode to create an import cycle
and edges are removed to make the graph acyclic.
%% \QALAN{}: do all modes remove the edges?

@@

Telemetry allows an application to ``phone home'' with data
summarizing usage patterns, such performance data, crash reporting,
and feature uptake. A typical usage is in deciding whether an API can
be deprecated; without telemetry, it may be difficult to know how
popular the API is, but with telemetry it is straightforward.
Telemetry for programming languages can, however, be controversial.
See, for example, the lively discussion around telemetry in the Go
toolchain~\cite{golang-telemetry}.

Roblox Studio has a framework for reporting telemetry.
Various features in the IDE use telemetry to measure their effectiveness.
The Luau open source toolchain does \emph{not} report telemetry,
as it is designed to be used in build environments such as Continuous Integration
servers, where hermetic deterministic builds are expected.

We added a telemetry subsystem to Roblox Studio in order to study type errors
in Luau code.
This section explains the constraints our telemetry had to meet~(\cref{s:tel-limitations})
and the final design~(\cref{s:telemetry-records}).

\subsection{Limitations for Luau Telemetry}
\label{s:tel-limitations}

Telemetry in Roblox Studio faces three major constraints:
\begin{itemize}

  \item
    It \emph{must not reveal private information}.
    This can include Personally Identifying Information (PII) about creators: their identity, location, etc.
    It also includes trade secrets. Even an error message that contains the
    name of a function can reveal something the creator does not want to share.

  \item
    On the client side, telemetry computations \emph{must run quickly enough} to avoid
    slowing down the editing experience.

  \item
    Telemetry \emph{must transmit a small amount of data}---not too many records
    with not too much information inside---to avoid overloading the servers.

\end{itemize}
Cox~\cite{transparent-telemetry} discusses in detail the privacy implications
and tradeoffs in telemetry.

In addition, due to the architecture of Roblox Studio
there is some information that is not available to the
typechecker:
\begin{itemize}
  \item
    Lifecycle events including as save, run, quit, and publish.\footnote{Publish events
    cannot be reported at any rate because they are public, and could be used
    to match a telemetry session to a creator.}
  \item
    GUI state, in particular, whether the Script Analysis widget is visible.
\end{itemize}
Our type error telemetry has to meet all of the above constraints.

@@

\begin{figure}[t]\centering
  \begin{tabular}{|l|l|l|l|l|l|l|}\hline
    ID & Time & Mode & Reason & Size Info & Overall Counts & Edit Range Counts
    \\\toprule
    \multicolumn{6}{c}{\raisebox{.5\normalbaselineskip}[0pt][0pt]{$\underbrace{\hspace{9cm}}$}} &
    \multicolumn{1}{c}{\raisebox{.5\normalbaselineskip}[0pt][0pt]{$\underbrace{\hspace{3cm}}$}} \\[-0.5ex]
    \multicolumn{6}{c}{fixed-size} & \multicolumn{1}{c}{variable-size}
  \end{tabular}
  \caption{Structure of a telemetry record. There are 13 overall counts and up to 70 edit range counts in each record.}
  \label{f:tdata}
\end{figure}

The type error telemetry for a Roblox Studio session is gathered as a series of
records, all with the same \panon{} session identifier.
This allows us to correlate telemetry across a single
session, but not between sessions.
In order to avoid swamping the telemetry servers, the client reports on a subset
of telemetry events via two levels of uniform random sampling:
\begin{itemize}
  \item
    1\% of Roblox Studio sessions generate \emph{any} type error telemetry, and
  \item
    0.5\% of all type analyses (approximately 1 in 200 keystrokes)
    in an enrolled session generate a telemetry record.
\end{itemize}
In addition to sending telemetry for randomly-selected keystrokes,
our system sends a telemetry record every time the session changes
focus from one module to another.
These records provide important context for mode switches:
if one record uses \mstrict{} mode but the next uses \mnocheck{},
then it is critical to know whether the creator downgraded modes
or simply switched to another module.

@@

To track the edit range, we record a start and end position, which we
update appropriately on every edit. This can result in very large and
imprecise edit ranges, for example, if the user edits at the beginning and end
of the file.
The upshot of this strategy is that it reduces the size and complexity of
telemetry records because there is only one inerval for errors to overlap with.

This telemetry design is admittedly coarse-grained.
For instance, it does not distinguish between edits that ignore a type
error from edits that remove one error while introducing another.
We acknowledge this and other threats in~\cref{s:threats}.
The main advantages of this telemetry are the small size of each record
and complete the lack of private information.
Furthermore, despite the limitations, this telemetry supports a variety of
interesting conclusions that we showcase in the next section.

@@

  \item
    \FILL{} beware counts include module switches. Why???
    Is that defensible at all?

@@

\FILL{} cannot match sessions to codebases

@@

In other professional-grade IDEs, such as VSCode~\cite{vscode-telemetry},
IntelliJ~\cite{intellij-telemetry}, and \code{.NET}~\cite{dotnet-telemetry}, telemetry may include all sorts of
data: error reports, source code fragments, timestamps, and filepaths.
Users may be able to opt out of some telemetry, but the details depend on the
license agreement.
Furthermore, at least in VSCode, IDE extensions can report their own telemetry.
While this data can be invaluable for discovering bugs in production, it
must be handled with extreme care.
